\section{The problem setup}

The objective of this project is to compute the minimum-time trajectory for a race car on a closed racing circuit. Unlike conventional path-following problems, where the trajectory is predefined, both the vehicle trajectory and the control inputs are optimized simultaneously while satisfying the physical limitations of the vehicle and the geometric constraints imposed by the track.

The optimization is performed on a real karting circuit located in Freiburg. The track geometry is represented by its centerline, local curvature and track width, which constitute the geometric input to the optimal control problem. These quantities are obtained by processing an image of the circuit and extracting its boundaries.

Besides the track geometry, the relevant vehicle parameters are assumed to be known. These include the wheelbase, steering angle limits, longitudinal acceleration and braking limits, and the maximum admissible lateral acceleration. Together, these parameters define the dynamic and physical constraints of the optimization problem.

The objective is therefore to determine the state and control trajectories that minimize the total lap time while ensuring that the vehicle remains within the track boundaries and satisfies all vehicle constraints.

To formulate this optimal control problem, an appropriate mathematical model of the vehicle is required. In the following section, a simplified kinematic bicycle model is introduced together with the assumptions adopted throughout this work.